---
title: "R Notebook"
output: html_notebook
---
Data example

Using geographic variation in college proximity to estimate the return to schooling

- Z: indicator that the subject grew up near a 4 year college
- A: subject's year of education
- Y: subject's income
- X: covariates such as paraents years of education, region of country, age, race, IQ score from a test in HS, etc

```{r}
library(ivpack)
```

```{r}
data(card.data)

# iv nearc4
# outcome lwage (log of wage)
# treatment educ

mean(card.data$nearc4)
par(mfrow=c(1,2))
hist(card.data$lwage)
hist(card.data$educ)
```

```{r}
```


```{r}
# make education binary
educ12<-card.data$educ>12

# estimate proportion of 'compliers'
propcomp = mean(educ12[card.data$nearc4==1])-mean(educ12[card.data$nearc4==0])

# itention to treat effect
itt<-mean(card.data$lwage[card.data$nearc4==1])-
  mean(card.data$lwage[card.data$nearc4==0])

# complier average causal effect
itt/propcomp
```


2SLS two stage least squares
```{r}
# stage1: regress A on Z
s1<-lm(educ12~card.data$nearc4)
## get predicted value of A given Z for each subject
predtx<-predict(s1, type="response")
table(predtx)

# stage2: regress y on predicted value of a
lm(card.data$lwage~predtx)
```

2SLS in ivpack
```{r}
ivmodel=ivreg(lwage~educ12, ~ nearc4, x=TRUE, data=card.data)
robust.se(ivmodel)
```

Controlling for covariates
```{r}
ivmodel=ivreg(lwage~educ12+exper+reg661+reg662+reg663+reg664+reg665+reg666+reg667+reg668, 
              ~ nearc4+exper+reg661+reg662+reg663+reg664+reg665+reg666+reg667+reg668, 
              x=TRUE, data=card.data)
robust.se(ivmodel)
```

