---
title: "R Notebook"
output: html_notebook
---
Data example: Right heart catheterization(心脏导管) data 

ICU patients in 5 hospitals
- Treatment: right heart catheterization vs not
- Outcome: death (yes/no)
- Confounders: demographics, insurance, disease diagnoses, etc
- 218 treated and 3551 controls

```{r}
library(tableone)
library(Matching)
```

```{r}
# read in data
load(url("http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/rhc.sav"))
```

```{r}
head(rhc)
```
```{r}
 str(rhc)
```


```{r}
ARF<-as.numeric(rhc$cat1=='ARF')# 急性呼吸衰竭
CHF<-as.numeric(rhc$cat1=='CHF') # 充血性心力衰竭
Cirr<-as.numeric(rhc$cat1=='Cirrhosis')# 肝硬化
colcan<-as.numeric(rhc$cat1=='Colon Cancer')# 结肠癌
Coma<-as.numeric(rhc$cat1=='Coma')# 昏迷
COPD<-as.numeric(rhc$cat1=='COPD')# 慢性阻塞性肺病
lungcan<-as.numeric(rhc$cat1=='Lung Cancer')# 肺癌
MOSF<-as.numeric(rhc$cat1=='MOSF w/Malignancy')# 恶性（肿瘤等）
sepsis<-as.numeric(rhc$cat1=='MOSF w/Sepsis')# 败血症
female<-as.numeric(rhc$sex=='Female')
died<-as.numeric(rhc$death=='Yes')
age<-rhc$age
treatment<-as.numeric(rhc$swang1=='RHC')
meanbp1<-rhc$meanbp1
```


```{r}
# covariates we will use(shorter list than you would use in practice)
xvars<-c("ARF", "CHF", "Cirr", "colcan", "Coma", 
         "lungcan", "MOSF", "sepsis", "age", "female", "meanbp1")
```

```{r}
# new dataset
mydata<-cbind(ARF, CHF, Cirr, colcan, Coma, lungcan, MOSF, 
              sepsis, age, female, meanbp1, treatment, died)
mydata<-data.frame(mydata)
```

```{r}
# creat a table 1, pre matching
table1 <- CreateTableOne(vars=xvars, strata="treatment", 
                         data=mydata, test=FALSE)
# include standardized mean difference(SMD)
print(table1, smd=TRUE)
```

```{r}
# do greedy matching on Mahalanobis distance
greedymatch<-Match(Tr=treatment, M=1, X=mydata[xvars])
matched<-mydata[unlist(greedymatch[c("index.treated", "index.control")]),]
```

```{r}
# creat a table 1, matched
matchedtable1 <- CreateTableOne(vars=xvars, strata="treatment", 
                         data=matched, test=FALSE)
# include standardized mean difference(SMD)
print(matchedtable1, smd=TRUE)
```

Outcome analysis: 
```{r}
# outcome analysis
y_trt<-matched$died[matched$treatment==1]
y_con<-matched$died[matched$treatment==0]
```

if we want a causal risk difference, we can carry out a paired t-test

- Point estimate: 0.045
  - Difference in probability of death if everyone received RHC versus if no one received RHC is 0.045, higher risk of death in RHC group
- 95% CI: (0.019, 0.072)
- P value: < 0.001

```{r}
# 方法1
# pairwise difference
diffy<-y_trt-y_con

# paired ttest
t.test(diffy)
```

McNemar test

```{r}
# 方法2
table(y_trt, y_con)
mcnemar.test(matrix(c(994, 493, 394, 305),2,2))
```

