---
title: "R Notebook"
output: html_notebook
---
Data example: Right heart catheterization(心脏导管) data 

ICU patients in 5 hospitals
- Treatment: right heart catheterization vs not
- Outcome: death (yes/no)
- Confounders: demographics, insurance, disease diagnoses, etc
- 218 treated and 3551 controls

```{r}
library(tableone)
library(MatchIt)
```

```{r}
# read in data
load(url("http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/rhc.sav"))
```

```{r}
head(rhc)
```

```{r}
ARF<-as.numeric(rhc$cat1=='ARF')# 急性呼吸衰竭
CHF<-as.numeric(rhc$cat1=='CHF') # 充血性心力衰竭
Cirr<-as.numeric(rhc$cat1=='Cirrhosis')# 肝硬化
colcan<-as.numeric(rhc$cat1=='Colon Cancer')# 结肠癌
Coma<-as.numeric(rhc$cat1=='Coma')# 昏迷
COPD<-as.numeric(rhc$cat1=='COPD')# 慢性阻塞性肺病
lungcan<-as.numeric(rhc$cat1=='Lung Cancer')# 肺癌
MOSF<-as.numeric(rhc$cat1=='MOSF w/Malignancy')# 恶性（肿瘤等）
sepsis<-as.numeric(rhc$cat1=='MOSF w/Sepsis')# 败血症
female<-as.numeric(rhc$sex=='Female')
died<-as.numeric(rhc$death=='Yes')
age<-rhc$age
treatment<-as.numeric(rhc$swang1=='RHC')
meanbp1<-rhc$meanbp1
aps<-rhc$aps1
```

```{r}
# new dataset
mydata<-cbind(ARF, CHF, Cirr, colcan, Coma, lungcan, MOSF, 
              sepsis, age, female, meanbp1, treatment, died, aps)
mydata<-data.frame(mydata)
```

Fit propensity score model: logistic regression 

```{r}
# logistic regression
psmodel<-glm(treatment~ARF+CHF+Cirr+colcan+Coma+lungcan+MOSF+
               sepsis+age+female+meanbp1+aps,
             family = binomial(), data=mydata)

# show coefficients ect
summary(psmodel)
# create propensity score
pscore<-psmodel$fitted.values
```

PS matching without caliper
```{r}
# do greedy matching on logti(ps)
psmatch<-Match(Tr=mydata$treatment, M=1, X=qlogis(pscore), replace=FALSE)
matched<-mydata[unlist(psmatch[c("index.treated", "index.control")]),]

xvars<-c("ARF", "CHF", "Cirr", "colcan", "Coma", 
         "lungcan", "MOSF", "sepsis", "age", "female", "meanbp1", "aps")

matchedtab1 <- CreateTableOne(vars=xvars, strata="treatment", 
                         data=matched, test=FALSE)
# include standardized mean difference(SMD)
print(matchedtab1, smd=TRUE)
```

Outcome analysis without caliper: 
```{r}
# outcome analysis
y_trt<-matched$died[matched$treatment==1]
y_con<-matched$died[matched$treatment==0]

# pairwise difference
diffy<-y_trt-y_con

# paired ttest
t.test(diffy)
```

PS matching with caliper
```{r}
# do greedy matching on logti(ps)
psmatch<-Match(Tr=mydata$treatment, M=1, X=qlogis(pscore), replace=FALSE, caliper=.2)
matched<-mydata[unlist(psmatch[c("index.treated", "index.control")]),]

xvars<-c("ARF", "CHF", "Cirr", "colcan", "Coma", 
         "lungcan", "MOSF", "sepsis", "age", "female", "meanbp1", "aps")

matchedtab1 <- CreateTableOne(vars=xvars, strata="treatment", 
                         data=matched, test=FALSE)
# include standardized mean difference(SMD)
print(matchedtab1, smd=TRUE)
```

Outcome analysis with caliper: 
```{r}
# outcome analysis
y_trt<-matched$died[matched$treatment==1]
y_con<-matched$died[matched$treatment==0]

# pairwise difference
diffy<-y_trt-y_con

# paired ttest
t.test(diffy)
```


R package: MatchIt
```{r}
# user matchit for propensity score, nearest neighbor matching
m.out<-matchit(treatment~ARF+CHF+Cirr+colcan+Coma+lungcan+MOSF+
               sepsis+age+female+meanbp1+aps,
             family = binomial(), data=mydata)
summary(m.out)

# propensity score plots
plot(m.out, type="jitter")
plot(m.out, type="hist")
```

